---
title: "In class Exercise 03"
execute: 
  warning: false # show warnings
  eval: true # evaluate the code
  echo: true # show the code
  freeze: true

date: "`r Sys.Date()`"
---

# Getting started!!!!

## install and load R packages

```{r}
pacman::p_load(tidyverse, sf, raster, spatstat, tmap,maptools)
```

## installing maptools

```{r}
#| echo: false
#| eval: false

install.packages("maptools", repos = "https://packagemanager.posit.co/cran/2023-10-13")
```

## importing spatial dataa

```{r}
childcare_sf <- st_read("data/ChildCareServices.geojson") %>%
  st_transform(crs = 3414)
```

```{r}
mpsz_sf <- st_read(dsn = "data", 
                layer = "MP14_SUBZONE_WEB_PL")
```

```{r}
sg_sf <- st_transform(st_union(mpsz_sf), crs = 3414)
```

```{r}
plot(sg_sf)
```

## store data in rds for future use

```{r}
write_rds(childcare_sf, "data/rds/childcare_sf.rds")
write_rds(mpsz_sf, "data/rds/mpsz_sf.rds")
write_rds(sg_sf, "data/rds/sg_sf.rds")
```

## mapping geospatial daa of childcare services

map mpsz_sf then add childcare_sf

```{r}
tm_shape(mpsz_sf)+
  tm_polygons()+
  tm_shape(childcare_sf)+
  tm_dots()
```

```{r}
tmap_mode('view')
tm_shape(childcare_sf)+
  tm_dots()
```

Remember to set back tmap mode to plot instead of view. avoid more than 10 interactive maps

```{r}
tmap_mode('plot')
```

## data wrangling

Converting sf data frames to sp’s Spatial\* class

```{r}
childcare <- as_Spatial(childcare_sf)
mpsz <- as_Spatial(mpsz_sf)
sg <- as_Spatial(sg_sf)
```

```{r}
childcare
```

```{r}
mpsz
```

```{r}
sg
```

## converting spatial classes into generic sp format

```{r}
childcare_sp <- as(childcare, "SpatialPoints")
sg_sp <- as(sg, "SpatialPolygons")
```

```{r}
childcare_sp
```

```{r}
sg_sp
```

## Converting the generic sp format into spatstat’s ppp format

```{r}
childcare_ppp <- as(childcare_sp, "ppp")
childcare_ppp
```

plot childcare ppp

```{r}
plot(childcare_ppp)
```

```{r}
summary(childcare_ppp)
```

check for duplicates

```{r}
any(duplicated(childcare_ppp))
```

count the number of co-indicence points

```{r}
multiplicity(childcare_ppp)

```

If we want to know how many locations have more than one point event, we can use the code chunk below.

```{r}
sum(multiplicity(childcare_ppp) > 1)
```

this shows that there are 338 duplicates

```{r}
tmap_mode('view')
tm_shape(childcare) +
  tm_dots(alpha=0.4, 
          size=0.05)
```

```{r}
tmap_mode('plot')
```

There are three ways to overcome this problem. The easiest way is to delete the duplicates. But, that will also mean that some useful point events will be lost.

The second solution is use jittering, which will add a small perturbation to the duplicate points so that they do not occupy the exact same space.

The third solution is to make each point “unique” and then attach the duplicates of the points to the patterns as marks, as attributes of the points. Then you would need analytical techniques that take into account these marks.

jittering approach

```{r}
childcare_ppp_jit <- rjitter(childcare_ppp, 
                             retry=TRUE, 
                             nsim=1, 
                             drop=TRUE)
```

check again

```{r}
any(duplicated(childcare_ppp_jit))
```

## creating owin object

When analysing spatial point patterns, it is a good practice to confine the analysis with a geographical area like Singapore boundary. In spatstat, an object called owin is specially designed to represent this polygonal region.

The code chunk below is used to covert sg SpatialPolygon object into owin object of spatstat.

```{r}
sg_owin <- as(sg_sp, "owin")
```

```{r}
plot(sg_owin)
```

```{r}
summary(sg_owin)
```

## Combining point events object and owin object

```{r}
childcareSG_ppp = childcare_ppp[sg_owin]
```

```{r}
summary(childcareSG_ppp)
```

```{r}
plot(childcareSG_ppp)
```

# First-order Spatial Point Patterns Analysis

## Kernel Density Estimation

```{r}
kde_childcareSG_bw <- density(childcareSG_ppp,
                              sigma=bw.diggle,
                              edge=TRUE,
                            kernel="gaussian") 
```

```{r}
plot(kde_childcareSG_bw)
```

retrieving bandwidth

```{r}
bw <- bw.diggle(childcareSG_ppp)
bw
```

rescalling kde values to change the legend or indicator on the right

```{r}
childcareSG_ppp.km <- rescale(childcareSG_ppp, 1000, "km")
```

```{r}
kde_childcareSG.bw <- density(childcareSG_ppp.km, sigma=bw.diggle, edge=TRUE, kernel="gaussian")
plot(kde_childcareSG.bw)
```

Working with different automatic badwidth methods

```{r}
 bw.CvL(childcareSG_ppp.km)
```

```{r}
bw.scott(childcareSG_ppp.km)
```

```{r}
bw.ppl(childcareSG_ppp.km)
```

```{r}
bw.diggle(childcareSG_ppp.km)
```

compare output of diggle and ppl

```{r}
kde_childcareSG.ppl <- density(childcareSG_ppp.km, 
                               sigma=bw.ppl, 
                               edge=TRUE,
                               kernel="gaussian")
par(mfrow=c(1,2))
plot(kde_childcareSG.bw, main = "bw.diggle")
plot(kde_childcareSG.ppl, main = "bw.ppl")
```

## Working with different kernel methods

```{r}
par(mfrow=c(2,2))
plot(density(childcareSG_ppp.km, 
             sigma=bw.ppl, 
             edge=TRUE, 
             kernel="gaussian"), 
     main="Gaussian")
plot(density(childcareSG_ppp.km, 
             sigma=bw.ppl, 
             edge=TRUE, 
             kernel="epanechnikov"), 
     main="Epanechnikov")
plot(density(childcareSG_ppp.km, 
             sigma=bw.ppl, 
             edge=TRUE, 
             kernel="quartic"), 
     main="Quartic")
plot(density(childcareSG_ppp.km, 
             sigma=bw.ppl, 
             edge=TRUE, 
             kernel="disc"), 
     main="Disc")
```

# Fixed and Adaptive KDE

## Computing KDE by using fixed bandwidth

```{r}
kde_childcareSG_600 <- density(childcareSG_ppp.km, sigma=0.6, edge=TRUE, kernel="gaussian")
plot(kde_childcareSG_600)
```

## Computing KDE by using adaptive bandwidth

```{r}
kde_childcareSG_adaptive <- adaptive.density(childcareSG_ppp.km, method="kernel")
plot(kde_childcareSG_adaptive)
```

compare both

```{r}
par(mfrow=c(1,2))
plot(kde_childcareSG.bw, main = "Fixed bandwidth")
plot(kde_childcareSG_adaptive, main = "Adaptive bandwidth")
```

## convert KDE output into grid object

```{r}
gridded_kde_childcareSG_bw <- as.SpatialGridDataFrame.im(kde_childcareSG.bw)
spplot(gridded_kde_childcareSG_bw)
```

convert grid object into raster

```{r}
kde_childcareSG_bw_raster <- raster(gridded_kde_childcareSG_bw)
```

```{r}
kde_childcareSG_bw_raster
```

assigning projection system ???

```{r}
projection(kde_childcareSG_bw_raster) <- CRS("+init=EPSG:3414")
kde_childcareSG_bw_raster
```

## visualising output

```{r}
tm_shape(kde_childcareSG_bw_raster) + 
  tm_raster("v") +
  tm_layout(legend.position = c("right", "bottom"), frame = FALSE)
```

Comparing Spatial Point Patterns using KDE

```{r}
pg = mpsz[mpsz@data$PLN_AREA_N == "PUNGGOL",]
tm = mpsz[mpsz@data$PLN_AREA_N == "TAMPINES",]
ck = mpsz[mpsz@data$PLN_AREA_N == "CHOA CHU KANG",]
jw = mpsz[mpsz@data$PLN_AREA_N == "JURONG WEST",]
```

plotting target areas

```{r}
par(mfrow=c(2,2))
plot(pg, main = "Ponggol")
plot(tm, main = "Tampines")
plot(ck, main = "Choa Chu Kang")
plot(jw, main = "Jurong West")
```

Converting the spatial point data frame into generic sp format

```{r}
pg_sp = as(pg, "SpatialPolygons")
tm_sp = as(tm, "SpatialPolygons")
ck_sp = as(ck, "SpatialPolygons")
jw_sp = as(jw, "SpatialPolygons")
```

owin object

```{r}
pg_owin = as(pg_sp, "owin")
tm_owin = as(tm_sp, "owin")
ck_owin = as(ck_sp, "owin")
jw_owin = as(jw_sp, "owin")
```

Combining childcare points and the study area

```{r}
childcare_pg_ppp = childcare_ppp_jit[pg_owin]
childcare_tm_ppp = childcare_ppp_jit[tm_owin]
childcare_ck_ppp = childcare_ppp_jit[ck_owin]
childcare_jw_ppp = childcare_ppp_jit[jw_owin]
```

Next, rescale() function is used to trasnform the unit of measurement from metre to kilometre.

```{r}
childcare_pg_ppp.km = rescale(childcare_pg_ppp, 1000, "km")
childcare_tm_ppp.km = rescale(childcare_tm_ppp, 1000, "km")
childcare_ck_ppp.km = rescale(childcare_ck_ppp, 1000, "km")
childcare_jw_ppp.km = rescale(childcare_jw_ppp, 1000, "km")
```

The code chunk below is used to plot these four study areas and the locations of the childcare centres.

```{r}
par(mfrow=c(2,2))
plot(childcare_pg_ppp.km, main="Punggol")
plot(childcare_tm_ppp.km, main="Tampines")
plot(childcare_ck_ppp.km, main="Choa Chu Kang")
plot(childcare_jw_ppp.km, main="Jurong West")
```

computing KDE

```{r}
par(mfrow=c(2,2))
plot(density(childcare_pg_ppp.km, 
             sigma=bw.diggle, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Punggol")
plot(density(childcare_tm_ppp.km, 
             sigma=bw.diggle, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Tempines")
plot(density(childcare_ck_ppp.km, 
             sigma=bw.diggle, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Choa Chu Kang")
plot(density(childcare_jw_ppp.km, 
             sigma=bw.diggle, 
             edge=TRUE, 
             kernel="gaussian"),
     main="JUrong West")
```

computing bandwidth

```{r}
par(mfrow=c(2,2))
plot(density(childcare_ck_ppp.km, 
             sigma=0.25, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Chou Chu Kang")
plot(density(childcare_jw_ppp.km, 
             sigma=0.25, 
             edge=TRUE, 
             kernel="gaussian"),
     main="JUrong West")
plot(density(childcare_pg_ppp.km, 
             sigma=0.25, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Punggol")
plot(density(childcare_tm_ppp.km, 
             sigma=0.25, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Tampines")
```

# nearest neighbour analysis

The test hypotheses are:

Ho = The distribution of childcare services are randomly distributed.

H1= The distribution of childcare services are not randomly distributed.

The 95% confident interval will be used.

## Testing spatial point patterns using Clark and Evans Test

```{r}
clarkevans.test(childcareSG_ppp,
                correction="none",
                clipregion="sg_owin",
                alternative=c("clustered"),
                nsim=99)
```

conlusion: looking at p value of 2.2e-16, we can reject the null hypothesis and conclude that the distribution of childcare services are not randomly distributed.

## Clark and Evans Test: Choa Chu Kang planning area

```{r}
clarkevans.test(childcare_ck_ppp,
                correction="none",
                clipregion=NULL,
                alternative=c("two.sided"),
                nsim=999)
```

## Clark and Evans Test: Tampines planning area

```{r}
clarkevans.test(childcare_tm_ppp,
                correction="none",
                clipregion=NULL,
                alternative=c("two.sided"),
                nsim=999)
```

Conclusion: looking at p value of choa chu kang and tampines,for tampines, we can reject the null hypothesis and conclude that the distribution of childcare services are not randomly distributed. However, for choa chu kang, we cannot reject the null hypothesis and conclude that the distribution of childcare services are randomly distributed.
