---
title: "Hands-on Exercise 11b - Calibrating Spatial Interaction Models with R"
format:
  html:
    code-fold: true
    code-summary: "Show the code"

execute: 
  eval: true
  echo: true
  warning: false
  freeze: true

date: "`r Sys.Date()`"
---

# 1.0 Getting started

```{r}
pacman::p_load(tmap, sf, sp,
               performance, reshape2,
               ggpubr, tidyverse)
```

```{r}
mpsz <- st_read(dsn = "data/geospatial",
                   layer = "MPSZ-2019") %>%
  st_transform(crs = 3414)
```
```{r}
mpsz
```
## 1.1 converting from sf data.table to sp class

```{r}
mpsz_sp <- as(mpsz, "Spatial")
mpsz_sp
```

# 2.0 computing the distance matrix

```{r}
dist <- spDists(mpsz_sp, 
                longlat = FALSE)
```

::: callout-note 
The distance is calculated between 2 centroids of a pair of polygons because the distance matrix is symmetric.

This means that the distance from polygon A to polygon B is the same as the distance from polygon B to polygon A.
:::

```{r}
head(dist, n=c(10, 10))

```

## 2.1 labelling column and row headers of a distance matrix 

```{r}
sz_names <- mpsz$SUBZONE_C
```

```{r}
colnames(dist) <- paste0(sz_names)
rownames(dist) <- paste0(sz_names)
```

## 2.2 pivoting distance value by SUBZONE_C

```{r}
distPair <- melt(dist) %>%
  rename(dist = value)
head(distPair, 10)
```

## 2.3 updating intro-zonal distances 

```{r}
distPair %>%
  filter(dist > 0) %>%
  summary()

```

Next, a constant distance value of 50m is added into intra-zones distance.
```{r}
distPair$dist <- ifelse(distPair$dist == 0,
                        50, distPair$dist)
```

```{r}
distPair %>%
  summary()
```
```{r}
distPair <- distPair %>%
  rename(orig = Var1,
         dest = Var2)
```


```{r}
write_rds(distPair, "data/rds/distPair.rds") 
```

```{r}
distPair <- read_rds("data/rds/distPair.rds")
```

# 3.0 Preparing flow data 

```{r}
od_data_fii <- read_rds("data/rds/od_data_fii.rds")
```

Next, we will compute the total passenger trip between and within planning subzones by using the code chunk below. The output is all flow_data.

```{r}
flow_data <- od_data_fii %>%
  group_by(ORIGIN_SZ, DESTIN_SZ) %>% 
  summarize(TRIPS = sum(MORNING_PEAK)) 
```

```{r}
head(flow_data, 10)
```

## 3.1 seperating intra-flow from passenger volume df 

```{r}
flow_data$FlowNoIntra <- ifelse(
  flow_data$ORIGIN_SZ == flow_data$DESTIN_SZ, 
  0, flow_data$TRIPS)
flow_data$offset <- ifelse(
  flow_data$ORIGIN_SZ == flow_data$DESTIN_SZ, 
  0.000001, 1)
```

## 3.2 combing passenger volume data with distance value 

```{r}
flow_data$ORIGIN_SZ <- as.factor(flow_data$ORIGIN_SZ)
flow_data$DESTIN_SZ <- as.factor(flow_data$DESTIN_SZ)
```

```{r}
flow_data1 <- flow_data %>%
  left_join (distPair,
             by = c("ORIGIN_SZ" = "orig",
                    "DESTIN_SZ" = "dest"))
```

# 4.0 Preparing origin and destination attributes 

```{r}
pop <- read_csv("data/aspatial/pop.csv")
```

```{r}
pop <- pop %>%
  left_join(mpsz,
            by = c("PA" = "PLN_AREA_N",
                   "SZ" = "SUBZONE_N")) %>%
  select(1:6) %>%
  rename(SZ_NAME = SZ,
         SZ = SUBZONE_C)
```

```{r}
flow_data1 <- flow_data1 %>%
  left_join(pop,
            by = c(ORIGIN_SZ = "SZ")) %>%
  rename(ORIGIN_AGE7_12 = AGE7_12,
         ORIGIN_AGE13_24 = AGE13_24,
         ORIGIN_AGE25_64 = AGE25_64) %>%
  select(-c(PA, SZ_NAME))
```

```{r}
flow_data1 <- flow_data1 %>%
  left_join(pop,
            by = c(DESTIN_SZ = "SZ")) %>%
  rename(DESTIN_AGE7_12 = AGE7_12,
         DESTIN_AGE13_24 = AGE13_24,
         DESTIN_AGE25_64 = AGE25_64) %>%
  select(-c(PA, SZ_NAME))
```

```{r}
write_rds(flow_data1, "data/rds/SIM_data.rds")
```

# 5.0 Calibrating the spatial interaction model

```{r}
SIM_data <- read_rds("data/rds/SIM_data.rds")
```

## 5.1 visualising the dependent variable 

```{r}
ggplot(data = SIM_data,
       aes(x = TRIPS)) +
  geom_histogram()
```

Notice that the distribution is highly skewed and not resemble bell shape or also known as normal distribution.

Next, let us visualise the relation between the dependent variable and one of the key independent variable in Spatial Interaction Model, namely distance.


```{r}
ggplot(data = SIM_data,
       aes(x = dist,
           y = TRIPS)) +
  geom_point() +
  geom_smooth(method = lm)
```

Notice that their relationship hardly resemble linear relationship.

On the other hand, if we plot the scatter plot by using the log transformed version of both variables, we can see that their relationship is more resemble linear relationship.

```{r}
ggplot(data = SIM_data,
       aes(x = log(dist),
           y = log(TRIPS))) +
  geom_point() +
  geom_smooth(method = lm)
```

## 5.2 checking for variables with zero values 

```{r}
summary(SIM_data)
```

Replace zero values to 0.99.

```{r}
SIM_data$DESTIN_AGE7_12 <- ifelse(
  SIM_data$DESTIN_AGE7_12 == 0,
  0.99, SIM_data$DESTIN_AGE7_12)
SIM_data$DESTIN_AGE13_24 <- ifelse(
  SIM_data$DESTIN_AGE13_24 == 0,
  0.99, SIM_data$DESTIN_AGE13_24)
SIM_data$DESTIN_AGE25_64 <- ifelse(
  SIM_data$DESTIN_AGE25_64 == 0,
  0.99, SIM_data$DESTIN_AGE25_64)
SIM_data$ORIGIN_AGE7_12 <- ifelse(
  SIM_data$ORIGIN_AGE7_12 == 0,
  0.99, SIM_data$ORIGIN_AGE7_12)
SIM_data$ORIGIN_AGE13_24 <- ifelse(
  SIM_data$ORIGIN_AGE13_24 == 0,
  0.99, SIM_data$ORIGIN_AGE13_24)
SIM_data$ORIGIN_AGE25_64 <- ifelse(
  SIM_data$ORIGIN_AGE25_64 == 0,
  0.99, SIM_data$ORIGIN_AGE25_64)
```

```{r}
summary(SIM_data)
```

## 5.3 Unconstrained spatial interaction model 
::: call-out note

The glm function is used to estimate the parameters of the spatial interaction model. The model is estimated using the Poisson distribution with a log link function. The independent variables are the log-transformed origin and destination population and the log-transformed distance between the origin and destination. The model is estimated using the SIM_data dataset.
:::

```{r}
uncSIM <- glm(formula = TRIPS ~ 
                log(ORIGIN_AGE25_64) + 
                log(DESTIN_AGE25_64) +
                log(dist),
              family = poisson(link = "log"),
              data = SIM_data,
              na.action = na.exclude)
uncSIM

```
## 5.4 R-squared function 

In order to measure how much variation of the trips can be accounted by the model we will write a function to calculate R-Squared value as shown below.

```{r}
CalcRSquared <- function(observed,estimated){
  r <- cor(observed,estimated)
  R2 <- r^2
  R2
}
```

```{r}
CalcRSquared(uncSIM$data$TRIPS, uncSIM$fitted.values)
```

```{r}
r2_mcfadden(uncSIM)
```

## 5.5 Origin (production) constrained SIM 

In this section, we will fit an origin constrained SIM by using the code chunk below.

```{r}
orcSIM <- glm(formula = TRIPS ~ 
                 ORIGIN_SZ +
                 log(DESTIN_AGE25_64) +
                 log(dist),
              family = poisson(link = "log"),
              data = SIM_data,
              na.action = na.exclude)
summary(orcSIM)
```

```{r}
CalcRSquared(orcSIM$data$TRIPS, orcSIM$fitted.values)
```
## 5.6 Destination constrained SIM

```{r}
decSIM <- glm(formula = TRIPS ~ 
                DESTIN_SZ + 
                log(ORIGIN_AGE25_64) + 
                log(dist),
              family = poisson(link = "log"),
              data = SIM_data,
              na.action = na.exclude)
summary(decSIM)
```

```{r}
CalcRSquared(decSIM$data$TRIPS, decSIM$fitted.values)
```

## 5.7 Doubly constrained SIM

```{r}
dbcSIM <- glm(formula = TRIPS ~ 
                ORIGIN_SZ + 
                DESTIN_SZ + 
                log(dist),
              family = poisson(link = "log"),
              data = SIM_data,
              na.action = na.exclude)
summary(dbcSIM)
```

```{r}
CalcRSquared(dbcSIM$data$TRIPS, dbcSIM$fitted.values)
```

# 6.0 model comparison 

we will be using the compare_performance() function from the performance package to compare the performance of the models.

```{r}
model_list <- list(unconstrained=uncSIM,
                   originConstrained=orcSIM,
                   destinationConstrained=decSIM,
                   doublyConstrained=dbcSIM)
```
We'll compare the models based on the RMSE metric.

```{r}
compare_performance(model_list,
                    metrics = "RMSE")
```
## 6.1 visualising fitted values 

```{r}
df <- as.data.frame(uncSIM$fitted.values) %>%
  round(digits = 0)
```

```{r}
SIM_data <- SIM_data %>%
  cbind(df) %>%
  rename(uncTRIPS = "uncSIM$fitted.values")
```


```{r}
df <- as.data.frame(orcSIM$fitted.values) %>%
  round(digits = 0)
```

```{r}
SIM_data <- SIM_data %>%
  cbind(df) %>%
  rename(orcTRIPS = "orcSIM$fitted.values")
```

```{r}
df <- as.data.frame(decSIM$fitted.values) %>%
  round(digits = 0)
```

```{r}
SIM_data <- SIM_data %>%
  cbind(df) %>%
  rename(decTRIPS = "decSIM$fitted.values")
```

```{r}
df <- as.data.frame(dbcSIM$fitted.values) %>%
  round(digits = 0)
```

```{r}
SIM_data <- SIM_data %>%
  cbind(df) %>%
  rename(dbcTRIPS = "dbcSIM$fitted.values")
```


```{r}
unc_p <- ggplot(data = SIM_data,
                aes(x = uncTRIPS,
                    y = TRIPS)) +
  geom_point() +
  geom_smooth(method = lm)

orc_p <- ggplot(data = SIM_data,
                aes(x = orcTRIPS,
                    y = TRIPS)) +
  geom_point() +
  geom_smooth(method = lm)

dec_p <- ggplot(data = SIM_data,
                aes(x = decTRIPS,
                    y = TRIPS)) +
  geom_point() +
  geom_smooth(method = lm)

dbc_p <- ggplot(data = SIM_data,
                aes(x = dbcTRIPS,
                    y = TRIPS)) +
  geom_point() +
  geom_smooth(method = lm)
```

We will now put all the graphs into a single visual using the ggarrange() function from the ggpubr package.

```{r}
ggarrange(unc_p, orc_p, dec_p, dbc_p,
          ncol = 2,
          nrow = 2)
```


# 7.0 Conclusion

In this exercise, we have calibrated four spatial interaction models, namely unconstrained, origin constrained, destination constrained, and doubly constrained models. We have also compared the performance of the models based on the RMSE metric. The doubly constrained model has the best fit compared to the other models. 

We can see that the doubly constrained model has the best fit compared to the other models. From the visualisation, we can see that the fitted values are close to the observed values. This implies that the model is able to predict the number of trips between origin and destination subzones accurately. 








