---
title: "Hands-on Exercise 11a - Processing and Visualising Flow Data"
format:
  html:
    code-fold: true
    code-summary: "Show the code"

execute: 
  eval: true
  echo: true
  warning: false
  freeze: true

date: "`r Sys.Date()`"
---

# 1.0 Getting Started

```{r}
pacman::p_load(tmap, sf, DT, stplanr, tidyverse)
```

# 2.0 Data

```{r}
#| eval: false

odbus <- read_csv("data/aspatial/origin_destination_bus_202210.csv")
```

```{r}
#| eval: false

glimpse(odbus)
```

```{r}
#| eval: false

odbus$ORIGIN_PT_CODE <- as.factor(odbus$ORIGIN_PT_CODE)
odbus$DESTINATION_PT_CODE <- as.factor(odbus$DESTINATION_PT_CODE) 
```

## 2.1 Extracting the study data

```{r}
#| eval: false

odbus6_9 <- odbus %>%
  filter(DAY_TYPE == "WEEKDAY") %>%
  filter(TIME_PER_HOUR >= 6 &
           TIME_PER_HOUR <= 9) %>%
  group_by(ORIGIN_PT_CODE,
           DESTINATION_PT_CODE) %>%
  summarise(TRIPS = sum(TOTAL_TRIPS))
```

```{r}
#| eval: false

datatable(odbus6_9)
```

::: callout-note
The datatable function is from the DT package. It is used to display the data in a tabular format.
:::

```{r}
#| eval: False
write_rds(odbus6_9, "data/rds/odbus6_9.rds")
```

```{r}
odbus6_9 <- read_rds("data/rds/odbus6_9.rds")
```

## 2.2 Working with geospatial data

```{r}
busstop <- st_read(dsn = "data/geospatial",
                   layer = "BusStop") %>%
  st_transform(crs = 3414)
```

```{r}
mpsz <- st_read(dsn = "data/geospatial",
                   layer = "MPSZ-2019") %>%
  st_transform(crs = 3414)
```

# 3.0 Geospatial data wrangling

```{r}
busstop_mpsz <- st_intersection(busstop, mpsz) %>%
  select(BUS_STOP_N, SUBZONE_C) %>%
  st_drop_geometry()
```

```{r}
datatable(busstop_mpsz)
```

```{r}
write_rds(busstop_mpsz, "data/rds/busstop_mpsz.rds")  
```

```{r}
od_data <- left_join(odbus6_9 , busstop_mpsz,
            by = c("ORIGIN_PT_CODE" = "BUS_STOP_N")) %>%
  rename(ORIGIN_BS = ORIGIN_PT_CODE,
         ORIGIN_SZ = SUBZONE_C,
         DESTIN_BS = DESTINATION_PT_CODE)
```

Check for duplicated records

```{r}
duplicate <- od_data %>%
  group_by_all() %>%
  filter(n()>1) %>%
  ungroup()
```

```{r}
od_data <- unique(od_data)
```

```{r}
od_data <- left_join(od_data , busstop_mpsz,
            by = c("DESTIN_BS" = "BUS_STOP_N")) 
```

```{r}
duplicate <- od_data %>%
  group_by_all() %>%
  filter(n()>1) %>%
  ungroup()
```

```{r}
od_data <- unique(od_data)
```

```{r}
od_data <- od_data %>%
  rename(DESTIN_SZ = SUBZONE_C) %>%
  drop_na() %>%
  group_by(ORIGIN_SZ, DESTIN_SZ) %>%
  summarise(MORNING_PEAK = sum(TRIPS))
```

```{r}
write_rds(od_data, "data/rds/od_data_fii.rds")
```

```{r}
od_data <- read_rds("data/rds/od_data_fii.rds")
```

# 4.0 visualising spatial interaction

We will now use the stplanr package to prepare the desire line.

## 4.1 removing intra-zonal flows

```{r}
od_data_fij <- od_data[od_data$ORIGIN_SZ!=od_data$DESTIN_SZ,]
```

```{r}
write_rds(od_data_fij, "data/rds/od_data_fij.rds")
```

```{r}
od_data_fij <- read_rds("data/rds/od_data_fij.rds")
```

## 4.2 creating desire lines

```{r}
flowLine <- od2line(flow = od_data_fij, 
                    zones = mpsz,
                    zone_code = "SUBZONE_C")
```

```{r}
write_rds(flowLine, "data/rds/flowLine.rds")
```

```{r}
flowLine <- read_rds("data/rds/flowLine.rds")
```

## 4.3 visualising the desire lines

```{r}
tm_shape(mpsz) +
  tm_polygons() +
flowLine %>%  
tm_shape() +
  tm_lines(lwd = "MORNING_PEAK",
           style = "quantile",
           scale = c(0.1, 1, 3, 5, 7, 10),
           n = 6,
           alpha = 0.3)
```

::: callout-note
The rendering process take more time because of the transparency arguement (alpha)
:::

```{r}
tm_shape(mpsz) +
  tm_polygons() +
flowLine %>%  
  filter(MORNING_PEAK >= 5000) %>%
tm_shape() +
  tm_lines(lwd = "MORNING_PEAK",
           style = "quantile",
           scale = c(0.1, 1, 3, 5, 7, 10),
           n = 6,
           alpha = 0.3)
```



